Discrete Quantum Theories and Computing

### 2019-01-22 dissertation.tex
dissertation.lyx "Chapter 6 Further Discussion" and CV Updated...

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map
with a conventional quantum probability measure, and the states consistent
with the composite QIVPM collapses to a single state as the maximum
length of intervals in $\mathscr{I}$, $\left\Vert \mathscr{I}\right\Vert $,
shrinks to $0$. In contrast, the globally inconsistent QIVPM defined
in Table~\ref{tab:non-Born-QIVPM} has the least sharp range~$\mathscr{I}_{0}$
with $\left\Vert \mathscr{I}_{0}\right\Vert =1$. This suggests a
possibility that shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize general QIVPMs, and it is natural to ask
whether there is a short enough length~$\varepsilon$ such that QIVPMs
mapping to intervals shorter than $\varepsilon$ always have non-empty
cores.

@@

 &  & \frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)\le1-\frameL\left(\ket{\psi_{0}}\right)\le\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)+2\varepsilon\,.\\

@@

then ask what $\frameL$ would look like with a positive $\varepsilon$.

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea of how hard this question
could be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum
probability space is glued by a family of classical probability spaces.
This is like the situation that a manifold is glued by many local
coordinates. When each small piece has \emph{exactly} the same and
positive curvature, the Killing-Hopf theorem asserts this manifold
is a sphere, but little can we say even if the curvature has an only
small deviation from constant. A similar situation might happen when
approximating Gleason's theorem, but this time the whole space is
glued by ``local'' classical probability space defined by each orthonormal
basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$. When the sum of
$\frameL$, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$, is
\emph{exactly} the same and equal to $1$, Gleason's theorem asserts
that $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
need to know a little bit more, like its QIVPM is a composite function,
to deduce its global property.

@@

After we build the quantum interval-valued probability model, we might
want to know how powerful a quantum computer could be based on this
model. Since the conventional quantum circuit model manipulates the
probability amplitudes instead of the measured probability, a quantum
computing model above QIVPMs might either need to manipulate the states
in the core of a QIVPM or find a way to manipulate QIVPMs directly.
However, both strategies are not straightforward. On one hand, a QIVPM
might have an empty core which cannot be evolved over time as we proved
in Thm.~\ref{thm:Non-extensible-of-Gleason's}. On the other hand,
if we want to manipulate and compute QIVPMs directly for multi-qubit
algorithms, we need to glue the QIVPM for each qubit or subsystem
together to get the QIVPM of the whole system, and this is not straightforward
either.

@@

\begin{definition}[Quantum Interval-valued Probability Measures
over Finite Fields]\label{def:QIVPMFF}Consider a vector space~$\mathcal{H}$
of dimension~$D$ over the complexified field~$\Fpp$, its set of
events~$\events_{p^{2}}$ as defined in Def.~\ref{def:QPMFF}, and
a collection of intervals~$\mathscr{I}$. A quantum interval-valued
probability measure over finite field $\bar{\mu}\colon\events_{p^{2}}\rightarrow\mathscr{I}$
assigns an interval to each event~$P$ subject to $\bar{\mu}(\mathbb{0})=\imposs$,
$\bar{\mu}(\mathbb{1})=\necess$, $\bar{\mu}\left(\mathbb{1}-P\right)=\necess-\bar{\mu}\left(P\right)$,
and satisfying for each pair of \emph{commuting} projectors~$P_{0}$
and~$P_{1}$ with $P_{0}P_{1}=P_{1}P_{0}$, 
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.\label{eq:QuantumInterval-valuedProbability-Inclusion-Finite-Field}
\end{equation}
\end{definition}

>>>>

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, Thm.~\ref{thm:Finite-precision-Gleason}
only applies to the QIVPMs constructed by composing an interval map
with a conventional quantum probability measure, and the states consistent
with the composite QIVPM collapses to a single state as the maximum
length of intervals in $\mathscr{I}$, $\left\Vert \mathscr{I}\right\Vert $,
shrinks to $0$. In contrast, the globally inconsistent QIVPM defined
in Table~\ref{tab:non-Born-QIVPM} has the least sharp range~$\mathscr{I}_{0}$
with $\left\Vert \mathscr{I}_{0}\right\Vert =1$. This suggests a
possibility that shrinking the length $\left\Vert \mathscr{I}\right\Vert $
might help to regularize general QIVPMs, and it is natural to ask
whether there is a short enough length~$\varepsilon$ such that QIVPMs
mapping to intervals bounded by $\varepsilon$ always have non-empty
cores.

@@

 &  & \frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)\le1-\frameL\left(\ket{\psi_{0}}\right)\le\frameL\left(\ket{\psi_{1}}\right)+\frameL\left(\ket{\psi_{2}}\right)+2\varepsilon\\

@@

then ask how $\frameL$ would look like with a positive $\varepsilon$.

With different settings, whether there is an approximate version of
Gleason's theorem was asked by Sam Sanders in constructivenews on
2013~\cite{Sanders2013}, and there is no clear answer for his question
as far as I understand. To have an idea of how hard this question
could be, recall in Sec.~\ref{sec:fuzzy} we state that a quantum
probability space is glued by a family of classical probability spaces.
\todo{Check the differential geometry analog with Andy, since I don't
really understand the differential geometry\ldots{} }This is like
the situation that a manifold is glued by many local coordinates.
When each small piece has \emph{exactly} the same and positive curvature,
the Killing-Hopf theorem asserts this manifold is a sphere, but little
can we say even if the curvature has a small deviation from constant.
A similar situation might happen when approximating Gleason's theorem,
but this time the whole space is glued by ``local'' classical probability
space defined by each orthonormal basis $\left\{ \ket{\psi_{i}}\right\} _{i=0}^{2}$.
When the sum of $\frameL$, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$,
is \emph{exactly} the same and equal to $1$, Gleason's theorem asserts
that $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
need to know a little bit more, like its QIVPM is a composite function,
to deduce its global property.

@@

After we build the quantum interval-valued probability model, we might
want to know how powerful a quantum computer could be based on this
model. Since the conventional quantum circuit model manipulates the
probability amplitudes instead of the measured probability, either
a quantum computing model above QIVPMs needs to manipulate the states
in the core of a QIVPM, or we need to find a way to manipulate QIVPMs
directly. However, both strategies are not straightforward. On one
hand, as we proved in Thm.~\ref{thm:Non-extensible-of-Gleason's},
a QIVPM might have an empty core which cannot be evolved over time.
On the other hand, if we want to manipulate and compute QIVPMs directly
for multi-qubit algorithms, we need to glue the QIVPM for each qubit
or subsystem together to get the QIVPM of the whole system, and this
is not straightforward either.

@@

\begin{definition}[Quantum Interval-valued Probability Measures
over Finite Fields]\label{def:QIVPMFF}Consider a vector space~$\mathcal{H}$
of dimension~$D$ over the complexified field~$\Fpp$, its set of
events~$\events_{p^{2}}$ as defined in Def.~\ref{def:QPMFF}, and
a collection of intervals~$\mathscr{I}$. A quantum interval-valued
probability measure over finite field $\bar{\mu}\colon\events_{p^{2}}\rightarrow\mathscr{I}$
assigns an interval to each event~$P$ subject to $\bar{\mu}(\mathbb{0})=\imposs$,
$\bar{\mu}(\mathbb{1})=\necess$, $\bar{\mu}\left(\mathbb{1}-P\right)=\necess-\bar{\mu}\left(P\right)$,
and satisfying for each pair of \emph{commuting} events~$P_{0}$
and~$P_{1}$ with $P_{0}P_{1}=P_{1}P_{0}$, 
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\,.\label{eq:QuantumInterval-valuedProbability-Inclusion-Finite-Field}
\end{equation}
\end{definition}
